---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mvbcf

<!-- badges: start -->
<!-- badges: end -->

Welcome to the mvbcf R package! This R package implements the Multivariate Bayesian Causal Forest model introduced in McJames et al., 2024:

McJames, N., O’Shea, A., Goh, Y. C. & Parnell, A. (2024). Bayesian causal forests for multivariate outcomes: Application to Irish data from an international large scale education assessment. *Journal of the Royal Statistical Society Series A: Statistics in Society*, 1-23. https://doi.org/10.1093/jrsssa/qnae049.

See below for details on installation. For worked examples and quick-start guides showing how to use the package, check out the vignettes and other materials. For questions, suggestions, or other queries, please reach out to me at mcjamesnathan\@yahoo.ie.

## Installation

You can install the development version of mvbcf with the following:

```{r, eval=FALSE}

if (!require("devtools")) {
  install.packages("devtools")
}
install_github("Nathan-McJames/mvbcf")
#Or to install vignettes with the package as well:
#install_github("Nathan-McJames/mvbcf", build_vignettes = TRUE)

```


